---
title: "WUP_04"
author: "Angelica Lopez Esparza"
date: "Summer 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
thedata<-read.csv(file="GSScsv.csv")
```
```{r eval=FALSE}
thedata <- thedata %>%
mutate(across(where(is.character), as.factor))
``` 


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE    

# Methods  
  
<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->


we wanted to find out by looking at the data what sort of data we had. we copied write up 3 to read the data in and changed the character variables into factor variables and saw that we had a numerical data which was WORDSUM and we now have a factor variable called degree.     
  

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
  
## Descriptive Results  

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->  

```{r}
library(plotly)

# Create the fancy boxplot
plot_ly(thedata, x = ~DEGREE, y = ~WORDSUM, type = "box") %>%
  layout(title = list(text = "<b>Fancy Boxplot</b>", x = 0.5, xanchor = "center", font = list(size = 14)),
         xaxis = list(title = "DEGREE"),
         yaxis = list(title = "WORDSUM"))

```  

I have five categories called BACHELOR, GARDUATE, HIGH SCHOOL, JR COLLEGE, and LT HIGH SCHOOL. from the graph i can see that the LT HIGH SCHOOL and GRADUATE are different. as well as LT HIGH SCHOOL and JR COLLEGE. if put back to back these boxes would not touch, which shows me that there is a difference so far between these few things.     

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
favstats(WORDSUM~DEGREE,data=thedata)
```

the average mean score for graduates is a bit over ten. there is a difference between the graduates and jr college average scores. further proving that the schoolings are different from each other  

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
model1 <- aov(WORDSUM~DEGREE,data=thedata)
summary.aov(model1)
```

there is a significant difference between the schooling of the five degrees. 

```{r}

# Perform one-way ANOVA
model <- aov(WORDSUM ~ DEGREE, data = thedata)

# Conduct post-hoc test (Tukey's HSD)
posthoc <- TukeyHSD(model)

# Print the results
print(posthoc)
```

based on the data given to us, LT HIGH SCHOOL and GRADUATES are different. high school and graduate are also different. each set comes with different levels. 


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
